Land-use legacies and climate change as a double challenge to oak forest resilience: mismatches of geographical and ecological rear edges

### 2020-06-10 man/ecosystems/ms_ecosystems.Rmd
update abstract

[^\*^]: Author Contributions. AJPL, GGI and RZ conceived of the study, conducted field work, and collected the data. AJPL and GGI performed the lab work. AJPL analyzed data and led the writing of the paper. GGI and RZ contributed in the writing process. All authors contributed to the drafts and gave final approval for publication.

@@

* The high values of resilience observed in our study suggest that *Q. pyrenaica* populations in Sierra Nevada are located in a geographical but not a climatic and ecological rear edge. Resilience of oak stands to drought events was not spatially homogeneous across the mountain range, due to differences in ecological conditions and/or past-management legacies. This is particularly relevant for rear-edge populations where topographic and biophysical variability can facilitate the existence of refugia.

@@

### Statistical analysis 
The severe drought events since 1901 were identified using SPEI-12 and regional climatic data (Figure 2). They were characterized in terms of duration, severity, intensity [see @Spinoni2015]. In a first step we explored temporal trends of EVI and BAI variables. Temporal trends of $EVI_{mean}$ were examined at the pixel scale, using the Mann–Kendall nonparametric test. Temporal trends for BAI were assessed using mean BAI chronologies. The impact of drought in greenness and growth was exploring using the $EVI_{sa}$ and the mean RWI site chronologies. Additionally the relationships between climatic variables and tree-growth variables (RWI and BAI site chronologies) were assessed using bootstrapped Pearson's correlations estimated using treeclim` [@Zang2015]. The non-climatic disturbance impacts on tree-growth were evaluated using the site disturbance chronologies.

>>>>

[^\*^]: Author Contributions. AJPL, GGI and RZ conceived of the study, conducted field work, and collected the data. AJPL and GGI performed the lab work. AJPL analyzed data and led the writing of the paper. All authors contributed in the writing process to the drafts and gave final approval for publication.

@@

* The high values of resilience observed suggest that the studied *Q. pyrenaica* populations are located in a geographical but not a climatic or ecological rear edge. Resilience of oak stands to drought events was not spatially homogeneous across the mountain range, due to differences in ecological conditions and/or past-management legacies. This is particularly relevant for rear-edge populations where topographic and biophysical variability can facilitate the existence of refugia.

@@

### Statistical analysis 
The severe drought events since 1901 were identified using SPEI-12 and regional climatic data (Figure 2). They were characterized in terms of duration, severity, intensity [see @Spinoni2015]. In a first step we explored temporal trends of EVI and BAI variables. Temporal trends of $EVI_{mean}$ were examined at the pixel scale, using the Mann–Kendall nonparametric test. Temporal trends for BAI were assessed using mean BAI chronologies. The impact of drought in greenness and growth was explored using the $EVI_{sa}$ and the mean RWI site chronologies. Additionally the relationships between climatic variables and tree-growth variables (RWI and BAI site chronologies) were assessed using bootstrapped Pearson's correlations estimated using treeclim` [@Zang2015]. The non-climatic disturbance impacts on tree-growth were evaluated using the site disturbance chronologies.
